---
title: "Lab 3 Homework"
author: "Anthony Weidner"
date: "`r Sys.Date()`"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Mammals Sleep
1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R.
```{r}
?msleep
```

The built-in data on mammal sleep patterns is based off V.M. Savage and G.B. West's study on a quantitative, theoretical framework for understanding mammalian sleep (Proceedings of the National Academy of Sciences, 2007). This information is from performing the above command. 

2. Store these data into a new data frame `sleep`.
```{r}
sleep <- msleep
sleep
```

3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.  
```{r}
dim(sleep)
str(sleep)
glimpse(sleep)
```

As shown by the code above, I used the commands dim(sleep), str(sleep), and glimpse(sleep) to derive more information about this data frame. The dimensions of the data frame are 83 rows by 11 columns (83 11). We know this because when looking at the actual data frame sleep (displayed in Question 2), it explicitly states that there are 83 rows and 11 variables, which is further confirmed by any of the three commands above. For example, the glimpse command tells us that are specifically 83 rows and 11 columns. We can see in the actual data frame that there are 83 observations/observed mammals (Dog, Cow, Owl, etc.) and 11 different variables (Genus, Vore, sleep_total, etc). This agrees with our stated dimensions.

4. Are there any NAs in the data? How did you determine this? Please show your code.  
```{r}
anyNA(sleep)
```

Yes, there are NAs in the data. I determined this by performing the anyNA(sleep) command to search for NA values in the data frame. Since the code returned TRUE, we have confirmed that there are NAs in the data.

5. Show a list of the column names is this data frame.
```{r}
names(sleep)
```


6. How many herbivores are represented in the data?  
```{r}
herbi_count <- subset(sleep,vore == "herbi")
herbi_count
count(herbi_count)
```

There are 32 herbivores represented in the data. When displaying herbi_count, we see that there are 32 rows of herbivores. We can confirm this count by doing count(herbi_count), which returns 32. 

7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
small <- subset(sleep,bodywt<=1)
small

large <- subset(sleep,bodywt>=200)
large

```


8. What is the mean weight for both the small and large mammals?
```{r}
mean(small$bodywt)
```

The mean weight for the small mammal group is 0.2596667 kg. 

```{r}
mean(large$bodywt)
```

The mean weight for the large mammal group is 1747.071 kg. 

9. Using a similar approach as above, do large or small animals sleep longer on average?  
```{r}
mean(small$sleep_total)
```

```{r}
mean(large$sleep_total)
```

According to the data, small animals sleep longer on average than large animals, with the average total sleep time for small animals being 12.65833 hours, in contrast with the average total sleep time for large animals being 3.3 hours. 

10. Which animal is the sleepiest among the entire dataframe?
```{r}
max(sleep$sleep_total)
```

```{r}
sleepy_mammal <- subset(sleep, sleep_total == 19.9)
sleepy_mammal
```

The little brown bat is the sleepiest animal among the entire data frame.

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   