---
title: "Lab 7 Homework"
author: "Anthony Weidner"
date: "`r Sys.Date()`"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
options(dplyr.summarise.inform = FALSE)
``` 

## Data
**1. For this homework, we will use two different data sets. Please load `amniota` and `amphibio`.**  

`amniota` data:  
Myhrvold N, Baldridge E, Chan B, Sivam D, Freeman DL, Ernest SKM (2015). “An amniote life-history
database to perform comparative analyses with birds, mammals, and reptiles.” _Ecology_, *96*, 3109.
doi: 10.1890/15-0846.1 (URL: https://doi.org/10.1890/15-0846.1).
```{r}
amniota <- readr::read_csv("data/amniota.csv")
amniota <- janitor::clean_names(amniota)
amniota
```

`amphibio` data:  
Oliveira BF, São-Pedro VA, Santos-Barrera G, Penone C, Costa GC (2017). “AmphiBIO, a global database
for amphibian ecological traits.” _Scientific Data_, *4*, 170123. doi: 10.1038/sdata.2017.123 (URL:
https://doi.org/10.1038/sdata.2017.123).
```{r}
amphibio <- readr::read_csv("data/amphibio.csv")
amphibio <- janitor::clean_names(amphibio)
amphibio
```

## Questions  
**2. Do some exploratory analysis of the `amniota` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  

```{r}
skim(amniota)
summary(amniota)
glimpse(amniota)
anyNA(amniota)
```

As shown by the analyses above, NAs are mostly represented by values of -999, and a handful of values in the female_maturity_d category have values of -30258.711, which should be treated as NAs. Note that there is no data with the value of NA.

**3. Do some exploratory analysis of the `amphibio` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  

```{r}
skim(amphibio)
summary(amphibio)
glimpse(amphibio)
anyNA(amphibio)
```

As shown by the analyses above, NAs are represented by the value of NA. 

**4. How many total NA's are in each data set? Do these values make sense? Are NA's represented by values?**   

```{r}
num_na_amniota = sum(is.na(amniota))
num_na_amniota
```

According to the command above, there are no NAs in the amniota data set. This is because NAs are represented by -999 or -30258.711 in this data set, not by NA. Thus, it makes sense that it shows that there are 0 values with the value of NA, since they are all represnted by -999 or -30258.711. 

```{r}
num_na_amphibio = sum(is.na(amphibio))
num_na_amphibio
```

According to the command above, there are 170691 NAs total in the data set. This value makes sense, since in this data set, NAs are actually represented by the value of NA. Thus, the number we are shown is accurate.

**5. Make any necessary replacements in the data such that all NA's appear as "NA".**   
```{r}
amniota <- amniota %>% 
  na_if("-999") %>%
  na_if("-30258.711")
```

There were some values of -30258.711 in the female_maturity_d category, so I made them NA because this value doesn't make any sense. This is in addition to making values of -999 as NA. 

```{r}
summary(amniota)
```

**6. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amniota` data.**  
```{r}
naniar::miss_var_summary(amniota)
```

**7. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amphibio` data.**
```{r}
naniar::miss_var_summary(amphibio)
```

**8. For the `amniota` data, calculate the number of NAs in the `egg_mass_g` column sorted by taxonomic class; i.e. how many NA's are present in the `egg_mass_g` column in birds, mammals, and reptiles? Does this results make sense biologically? How do these results affect your interpretation of NA's?**  

```{r}
amniota %>%
  group_by(class) %>%
  select(class, egg_mass_g) %>% 
  naniar::miss_var_summary(order=T)
```

Yes, these results make sense biologically because the data for egg mass for Mammals is entirely NA (100% NA). This occurs because Mammals do not have eggs, so there cannot be any data for egg mass. These results made me realize that NA data isn't just for missing data - it's also for data that is simply not applicable or possible, such as "mammal egg mass." Thus, I now am more aware of how NAs can be very commonly present in data sets, and that it is important to know how to deal with them.

**9. The `amphibio` data have variables that classify species as fossorial (burrowing), terrestrial, aquatic, or arboreal.Calculate the number of NA's in each of these variables. Do you think that the authors intend us to think that there are NA's in these columns or could they represent something else? Explain.**
```{r}
amphibio %>%
  select(fos,ter,aqu,arb) %>%
  naniar::miss_var_summary(order=T)
```

I think that the authors intend us to think that the NAs in these columns represent that the species does not have the certain classification (fossorial, terrestrial, aquatic, arboreal). For example, if a species was not fossorial, the author would put a value of NA for this classification to represent this fact. A value of 1 in this case means that the species satisfies the classification. Thus, a value of NA for these columns does not really mean NA, but rather a NO for the given classification. 

**10. Now that we know how NA's are represented in the `amniota` data, how would you load the data such that the values which represent NA's are automatically converted?**
```{r}
amniota <- 
  readr::read_csv(file = "data/amniota.csv", 
                  na = c("-30258.711", "-999"))
```

```{r}
summary(amniota)
```


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  